Parallel Quadtree Construction on Collections of Objects

### 2017-05-06 paper.tex
Layout changes

We refer to quadtree leaf cells that intersect two or more objects as ``conflict cells.'' A necessary and sufficient condition for a quadtree to resolve objects is to have no conflict cells. Our approach to computing such a quadtree is \blue{in two stages. We first build an initial quadtree, called the ``vertex quadtree,'' using a set $S$ of point samples. We initialize $S$ to be the object vertices. The second stage is to detect conflict cells in parallel, followed by augmenting $S$ with sample points such that a subsequent quadtree built on $S$ resolves conflict cells.} If $S$ changed, then we iterate (see section \ref{sec:iterate}) \blue{which is necessary only if a conflict cell has multiple intersecting objects and is minimized by starting from an initial vertex quadtree}. \blue{This two-stage approach enabled by our sample point computation is the first method that can resolve between objects fully in parallel regardless of object spacing, i.e., we do not iterate through levels of the quadtree, subdividing as we go.}

@@

%-------------------------------------------------------------------------------
% Implementation
%-------------------------------------------------------------------------------
\section{Implementation}
\blue{We have implemented\footnote{
Source code will be made available at our website. %is available at \url{http://cedmav.org/research/project/33-gvds.html}.
}
our algorithm using OpenCL. Figure \ref{fig:flow-chart} shows the stages of the major kernels. Every kernel call is parallel on vertices, facets, quadtree nodes or Morton code bits. Our implementation uses 64-bit Morton codes, which were sufficient for all of our datasets. There is no way of knowing \textit{a priori} what the object spacing is going to be, however, 64-bit codes were sufficient for the most demanding datasets while retaining reasonable radix-sort timings.
Choosing a large number of bits for the Morton code results in little or no wasted effort in later refinement. Increasing the size of the Morton codes affects the amount of later pruning only if it results in extraneous resolution of vertices in the initial vertex quadtree computation. The number of bits also has no effect on the number of conflict cells unless the Morton codes are too small to resolve vertices.}

\blue{Our implementation of the algorithm supports polygons and polylines which needn't be manifold or connected. Intersecting lines are not handled as a special case, i.e., the quadtree is simply resolved to its maximum depth.}

@@

\begin{figure}
  \centering
  \subfloat[][]{
    \label{fig:vascular-1}
    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/04-Pruned.PNG} }
  \subfloat[][]{
    \label{fig:vascular-2}
    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/06-Initial-Conflicts.PNG} }
  \subfloat[][]{
    \label{fig:vascular-3}
%    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/08-Resolved-Quadtree.PNG} }
    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/14-Resolved-Quadtree.PNG} }
  \caption{A large set of uniquely labeled polygons constructed from connected component analysis on a photograph of vascular cambium, a type of plant tissue.
    \protect\subref{fig:vascular-1} Initial vertex quadtree after pruning.
    \protect\subref{fig:vascular-2} All conflict cells of the initial quadtree.
    \protect\subref{fig:vascular-3} After conflict cell resolution. No quadtree cell intersects more than one object. Our method works even though objects in this dataset are often non-manifold and have self-intersections.
  }
  \label{fig:vascular}
\end{figure}

@@

%-------------------------------------------------------------------------------
% Conclusions
%-------------------------------------------------------------------------------
%\section{Conclusions}

>>>>

We refer to quadtree leaf cells that intersect two or more objects as ``conflict cells.'' A necessary and sufficient condition for a quadtree to resolve objects is to have no conflict cells. Our approach to computing such a quadtree is \blue{in two stages. We first build an initial quadtree, called the ``vertex quadtree,'' using a set $S$ of point samples. We initialize $S$ to be the object vertices. The second stage is to detect conflict cells in parallel, followed by augmenting $S$ with sample points such that a subsequent quadtree built on $S$ resolves conflict cells.} If $S$ changed, then we iterate (see section \ref{sec:iterate}) \blue{which is necessary only if a conflict cell has multiple intersecting objects. The number of iterations is minimized by starting from an initial vertex quadtree}. \blue{This two-stage approach enables us to resolve between objects fully in parallel regardless of object spacing, i.e., we do not iterate through levels of the quadtree, subdividing as we go.}

@@

%-------------------------------------------------------------------------------
% Implementation
%-------------------------------------------------------------------------------
\section{Implementation}
\blue{We have implemented\footnote{
Source code will be made available at our website. %is available at \url{http://cedmav.org/research/project/33-gvds.html}.
}
our algorithm using OpenCL. Figure \ref{fig:flow-chart} shows the stages of the major kernels. Every kernel call is parallel on vertices, facets, quadtree nodes or Morton code bits. Our implementation uses 64-bit Morton codes, which were sufficient for all of our datasets. There is no way of knowing \textit{a priori} what the object spacing is going to be, however, 64-bit codes were sufficient for the most demanding datasets while retaining reasonable radix-sort timings.
Choosing a large number of bits for the Morton code results in little or no wasted effort in later refinement, affecting the amount of later pruning only if intra-object vertex spacing is small compared to spacing between objects. The number of bits also has no effect on the number of conflict cells unless the Morton codes are too small to resolve vertices.}

\blue{Our implementation of the algorithm supports polygons and polylines which needn't be manifold or connected. Intersecting lines are not handled as a special case, i.e., the quadtree is simply resolved to its maximum depth. Special handling can be implemented per application as needed, e.g., for collision detection applications.}

@@

\begin{figure}[t]
  \centering
  \subfloat[][]{
    \label{fig:vascular-1}
    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/04-Pruned.PNG} }
  \subfloat[][]{
    \label{fig:vascular-2}
    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/06-Initial-Conflicts.PNG} }
  \subfloat[][]{
    \label{fig:vascular-3}
%    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/08-Resolved-Quadtree.PNG} }
    \includegraphics[width=0.31\columnwidth]{Nathan/vascular-bundles/14-Resolved-Quadtree.PNG} }
  \caption{A large set of uniquely labeled polygons constructed from connected component analysis on a photograph of vascular cambium, a type of plant tissue.
    \protect\subref{fig:vascular-1} Initial vertex quadtree after pruning.
    \protect\subref{fig:vascular-2} All conflict cells of the initial quadtree.
    \protect\subref{fig:vascular-3} After conflict cell resolution. No quadtree cell intersects more than one object. Our method works even though objects in this dataset are often non-manifold and have self-intersections.
  }
  \label{fig:vascular}
\end{figure}

@@
